---
title: "WiN Prep "
author: "Dani Cosme & Taurean Butler"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen=999)
```

# load packages
```{r}
if(!require('pacman')) {
	install.packages('pacman')
}

pacman::p_load(tidyverse, devtools, DescTools)

if (!require(scorequaltrics)) {
  devtools::install_github('dcosme/qualtrics', ref = "dev/enhance")
}
```

# load and tidy data {.tabset}
## load
```{r}
# non-demographic data
data_clean = read.csv("../../Data/Cleaned/pre_clean_long.csv", stringsAsFactors = FALSE) %>%
  bind_rows(read.csv("../../Data/Cleaned/daily_clean_long.csv", stringsAsFactors = FALSE)) %>%
  bind_rows(read.csv("../../Data/Cleaned/post_clean_long.csv", stringsAsFactors = FALSE)) %>%
  filter(!grepl("timing|timer", scale_name)) %>%
  rename("date" = StartDate) %>%
  select(survey_name, day, date, condition, SID, everything())

# demographic data
demo_clean = read.csv("../../Data/Cleaned/pre_clean_long_demo.csv", stringsAsFactors = FALSE)

# identity codes
identities = data.frame(item = as.character(1:12), item_text = c("military", "mental_health",  "disability", "ses", "race_ethnicity", "gender_sex", "sexuality", "age", "job_role", "education", "politics", "other"))
```

## daily items {.tabset}
### emotions
```{r}
emotions = data_clean %>% 
  filter(scale_name == "emotion") %>%
  mutate(item = dplyr::recode(item, "1" = "happy", 
                              "2" = "sad",
                              "3" = "angry", 
                              "4" = "anxious",
                              "5" = "stressed", 
                              "6" = "lonely",
                              "7" = "excited", 
                              "8" = "calm")) %>%
  arrange(SID, day)
```

### alcohol
Includes responses in the post survey

```{r}
alcohol_motivation = data_clean %>%
  filter(grepl("alcohol_motivation", scale_name)) %>%
  mutate(
    item = dplyr::recode(item,
                         "1" = "forget_worries",
                         "2" = "cheer_up",
                         "3" = "excitement",
                         "4" = "fun", 
                         "5" = "celebration",
                         "6" = "social",
                         "7" = "fit_in",
                         "alcohol_motivation_TEXT" = "other")) # fix as needed

alcohol = data_clean %>%
  filter(grepl("^alcohol_", scale_name)) %>%
  filter(!grepl("alcohol_motivation", scale_name)) %>%
  mutate(scale_name = "alcohol",
         item = gsub("alcohol_", "", item)) %>%
  bind_rows(alcohol_motivation) %>%
  arrange(SID, day)
```

### intervention-related items
```{r}
IIs = data_clean %>% 
  filter(grepl("II_", scale_name)) %>%
  filter(!condition == "control") %>%
  mutate(match = ifelse(scale_name == sprintf("II_%s", condition), "yes", "no"),
         item = dplyr::recode(item, "1" = "if", "2" = "then")) %>%
  filter(match == "yes") %>%
  select(-match)
  
study_peer = data_clean %>% 
  filter(!day == 0) %>%
  filter(scale_name == "strategy_use" | (scale_name == "study_peer" & grepl("perspective", condition))) 

daily_intervention = data_clean %>%
  filter(grepl("strategy_use", scale_name)) %>%
  bind_rows(., IIs)%>%
  bind_rows(., study_peer) %>%
  arrange(SID, day)
```

### interaction
```{r}
identity = data_clean %>%
  filter(grepl("int_identity_type|similarity_type", scale_name)) %>%
  left_join(., identities) %>%
  mutate(item_text = case_when((scale_name == "int_identity_type" & item == 14) | (grepl("similarity_type", scale_name) & item == 13) ~ "no interactions", .default = item_text),
         item = ifelse(!is.na(item_text), item_text, item),
         scale_name = gsub("int_", "interaction_", scale_name)) %>%
  select(-item_text)

partner = data_clean %>%
  filter(scale_name == "int_partner") %>%
  mutate(item = dplyr::recode(
    item,
    "1" = "family",
    "2" = "romantic",
    "3" = "friend",
    "4" = "work_colleague",
    "5" = "work_supervisor",
    "6" = "work_supervisee",
    "7" = "work_client",
    "8" = "work_mentorship",
    "9" = "acquaint",
    "10" = "stranger",
    "11" = "other",
    "other" = "other_text",
    "12" = "no interactions"),
  scale_name = gsub("int_", "interaction_", scale_name)
  )

outcome = data_clean %>%
  filter(scale_name == "int_outcome") %>%
  mutate(item = dplyr::recode(
    item,
    "1" = "felt_seen",
    "2" = "shared_understanding",
    "3" = "understood_other",
    "4" = "constructive_discussion",
    "5" = "felt_better"),
    scale_name = gsub("int_", "interaction_", scale_name)
    )

response = data_clean %>%
  filter(scale_name == "int_response") %>%
  mutate(item = dplyr::recode(
    item,
    "1" = "disclosure",
    "2" = "suppression",
    "3" = "reappraisal",
    "4" = "perspective_taking",
    "5" = "acceptance",
    "6" = "rumination",
    "7" = "self_monitoring"),
    scale_name = gsub("int_", "interaction_", scale_name)
    )

interaction = data_clean %>%
  filter(grepl("^int_", scale_name)) %>%
  filter(!grepl("identity_type|similarity_type|partner|outcome|response", scale_name)) %>%
  mutate(scale_name = "interaction",
         item = gsub("int_", "", item)) %>%
  bind_rows(., identity) %>%
  bind_rows(., outcome) %>%
  bind_rows(., partner) %>%
  bind_rows(., response) %>%
  arrange(SID, day)
```

### hassels & uplifts
```{r}
hassels_uplifts <- data_clean %>%
  filter(grepl("hassles|uplifts", scale_name)) %>%
  mutate(item = dplyr::recode(
    item,
    "1" = "close_relations",
    "2" = "workplace_relations",
    "3" = "distant_relations",
    "4" = "health_wellbeing",
    "5" = "physical_appearance",
    "6" = "work_education",
    "7" = "financial_legal",
    "8" = "news_politics",
    "9" = "home_living",
    "10" = "recreation_free_time",
    "11" = "spiritual_religious",
    "12" = "responsibilities_errands",
    "13" = "transportation",
    "14" = "other",
    "other" = "other_text",
    "15" = "none"
    )) %>%
  arrange(SID, day)
```

### well-being and sharing
```{r}
wellbeing = data_clean %>%
  filter(grepl("wellbeing|compassion|disclosure", scale_name)) %>%
  mutate(item = gsub("disclosures_", "", item),
         scale_name = gsub("_self|_other", "", scale_name),
         item = case_when(scale_name == "compassion" & item == "1" ~ "self",
                          scale_name == "compassion" & item == "2" ~ "others",
                          .default = item),
         item = dplyr::recode(
           item,
           "1" = "social_connection", 
           "2" = "purpose",
           "3" = "self_clarity", 
           "4" = "seen_understood", 
           "5" = "goal_progress", 
           "6" = "autonomous_motivation",
           "7" = "controlled_motivation"
          ),
         scale_name = case_when(grepl("connection|purpose|clarity|seen", item) ~ "wellbeing",
                                grepl("progress|motivation", item) ~ "goals_motivation",
                                .default = scale_name)) %>%
  arrange(SID, day)
```

### work
```{r}
work_identity = data_clean %>%
  filter(grepl("work_identity_type", scale_name)) %>%
  left_join(., identities) %>%
  mutate(item = ifelse(!is.na(item_text), item_text, item)) %>%
  select(-item_text)

work = data_clean %>%
  filter(grepl("work_attendance|work_experiences", scale_name)) %>%
  mutate(
    item = dplyr::recode(
    item,
    "1" = "belonging",
    "9" = "value",
    "10" = "reciprocity",
    "11" = "fit"
    ),
    scale_name = ifelse(!scale_name == "work_attendance", "work_VRF", scale_name)
  ) %>%
  bind_rows(., work_identity) %>%
  arrange(SID, day)
```

### merge daily items
```{r}
daily = emotions %>%
  bind_rows(., alcohol) %>%
  bind_rows(., daily_intervention) %>%
  bind_rows(., interaction) %>%
  bind_rows(., hassels_uplifts) %>%
  bind_rows(., wellbeing) %>%
  bind_rows(., work) %>%
  arrange(SID, day)

```

## pre and post items {.tabset}

### score data {.tabset}
```{r}
# get participant info for later merging
participant_info = data_clean %>%
  filter(!grepl("Daily", survey_name)) %>%
  select(survey_name, day, date, SID, condition) %>% 
  unique()

# specify rubric paths
scoring_rubrics = data.frame(file = dir(file.path("scoring_rubrics"), 
                                        pattern = '.*.csv',
                                        full.names = TRUE))
# read in rubrics
scoring_data = scorequaltrics::get_rubrics(scoring_rubrics,
                                                type = 'scoring')
```

#### DVS
Variables included at pre and post:

* Scales: AUDIT, CES-D-10, DERS-SF, ERQ, FS, GAD-7, IRQ, PSOM, ULS-8, VRF

```{r}
# subset df
data_prepped = data_clean %>%
  filter(grepl("AUDIT|CES|DERS|ERQ|FS|GAD|IRQ|PSOM$|ULS|VRF$", scale_name)) %>%
  mutate(item = ifelse(!scale_name == item, sprintf("%s_%s", scale_name, item), item),
         value = as.numeric(value)) %>%
  select(-survey_name) %>%
  dplyr::rename("survey_name" = scale_name)

# score
scored_pre = scorequaltrics::score_questionnaire(filter(data_prepped, day == 0),
                                                 scoring_data, SID = "SID", psych = FALSE)  %>%
  mutate(day = 0)

scored_post = scorequaltrics::score_questionnaire(filter(data_prepped, day == 13),
                                                 scoring_data, SID = "SID", psych = FALSE) %>%
  mutate(day = 13)

dvs_scored = scored_pre %>%
  bind_rows(scored_post) %>%
  mutate(score = as.numeric(score),
         score1 = case_when(scale_name == "PSOM" & n_missing == 18 ~ NaN, .default = score)) %>% #recode summed scores if missing all values
  dplyr::rename("value" = score,
                "item" = scored_scale) %>%
  select(-survey_name) %>%
  left_join(., participant_info) %>%
  select(survey_name, day, date, condition, SID, scale_name, item, value, n_missing) %>%
  arrange(SID, day)
```

#### individual differences

Variables included at pre only:

* Scales: DMQ-R-SF, SCC, SNI, VRF_demo, military_combat

* Single items: PSOM_arrangement, PSOM_level, military_branch, military_experience, military_date, politics_ideology, readiness_change, therapy

!! MISSING SCORING: SNI


```{r}
# subset df
data_prepped_ind = data_clean %>%
  filter(grepl("DMQ-R-SF|SCC|VRF_demo", scale_name)) %>%
  mutate(item = sprintf("%s_%s", scale_name, item),
         value = as.numeric(value)) %>%
  select(-survey_name) %>%
  dplyr::rename("survey_name" = scale_name)

# score
scored_pre_ind = scorequaltrics::score_questionnaire(filter(data_prepped_ind, day == 0),
                                                 scoring_data, SID = "SID", psych = FALSE)  %>%
  mutate(day = 0)

scored_ind_diffs = scored_pre_ind %>%
  #mutate(score = as.numeric(score)) %>%
  dplyr::rename("value" = score,
                "item" = scored_scale) %>%
  select(-survey_name) %>%
  left_join(., participant_info) %>%
  select(survey_name, day, date, condition, SID, scale_name, item, value, n_missing)

# score SNI
sni = data_clean %>%
  filter(grepl("SNI", scale_name)) 

# score military combat
military_combat = data_clean %>%
  filter(grepl("combat", scale_name)) %>%
  mutate(value = as.numeric(value),
         value = case_when(item == 1 ~ (value - 1) * 2,
                            item == 2 ~ (value - 1),
                            item == 3 ~ (value - 1) * 2,
                            item == 4 ~ (value - 1) * 2, .default = NA)) %>%
  group_by(survey_name, day, date, condition, SID, scale_name, flag_condition_missing) %>%
  summarize(value = sum(value))


# recode military identity and branch
military_identity = data_clean %>%
  filter(grepl("military_identity", scale_name)) %>%
  filter(value == "On") %>%
  mutate(value = item,
         item = "military_identity")

military_branch = data_clean %>%
  filter(grepl("military_branch", scale_name)) %>%
  mutate(item = case_when(item == "1" ~ "Army",
                          item == "2" ~ "Marines",
                          item == "3" ~ "Navy",
                          item == "4" ~ "Air Force",
                          item == "5" ~ "Space Force",
                          item == "6" ~ "Coast Guard",
                          item == "7" ~ "Other", .default = NA)) %>%
  filter(!is.na(value))

# recode PSOM arrangement and level items
PSOM_other = data_clean %>%
  filter(grepl("PSOM_", scale_name)) %>%
  mutate(item = case_when(scale_name == "PSOM_arrangement" & item == "1" ~ "In-person",
                          scale_name == "PSOM_arrangement" & item == "2" ~ "Hybrid",
                          scale_name == "PSOM_arrangement" & item == "3" ~ "Remote",
                          scale_name == "PSOM_arrangement" & item == "4" ~ "Self-employed", .default = item),
         value = case_when(scale_name == "PSOM_level" & value == "1" ~ "Entry-level",
                           scale_name == "PSOM_level" & value == "2" ~ "Individual contributor",
                           scale_name == "PSOM_level" & value == "3" ~ "Manager / supervisor",
                           scale_name == "PSOM_level" & value == "4" ~ "Director or above",
                           scale_name == "PSOM_level" & value == "5" ~ "Executive leadership", .default = value)) %>%
  filter(!(scale_name == "PSOM_arrangement" & is.na(value)))

# merge with single items
ind_diffs_single = data_clean %>%
  filter(grepl("military_date|military_experience|military_role|politics_ideology|readiness_change|therapy_", scale_name))

ind_diffs_merged = scored_ind_diffs %>%
  bind_rows(., ind_diffs_single) %>%
  bind_rows(., sni) %>%
  bind_rows(., military_identity) %>%
  bind_rows(., military_branch) %>%
  bind_rows(., PSOM_other) %>%
  arrange(SID, scale_name, item)
```

### intervention-related items

```{r}
intervention = data_clean %>%
  filter(grepl("alc_mind|alc_persp|dd_|gen_mind|gen_persp|study_peer", scale_name)) %>%
  filter(!(condition == "control" & (grepl("mind|persp|study_peer", scale_name) | grepl("mind|persp", item)))) %>%
  filter(!(condition == "gen_mindful" & (grepl("alc_mind|persp|study_peer", scale_name) | grepl("alc_mind|persp", item)))) %>%
  filter(!(condition == "alc_mindful" & (grepl("gen_mind|persp|study_peer", scale_name) | grepl("gen_mind|persp", item)))) %>%
  filter(!(condition == "gen_perspective" & (grepl("alc_persp|mind", scale_name) | grepl("alc_persp|mind", item)))) %>%
  filter(!(condition == "alc_perspective" & (grepl("gen_persp|mind", scale_name) | grepl("gen_persp|mind", item)))) %>%
  arrange(SID)
```

# data quality
```{r}
feedback = data_clean %>%
  filter(grepl("feedback", scale_name)) %>%
  arrange(SID, day)
```

# exclude participants
```{r}
# placeholder for additional exclusions
exclusions = data.frame()

# exclusions %>%
#   group_by(reason) %>%
#   dplyr::summarize(n = n())
```

Number of participants excluded = `r nrow(exclusions)`

# demographic data
```{r}
ses_subjective = demo_clean %>%
  filter(scale_name == "ses_subjective" & value == "On") %>%
  select(-value) %>%
  dplyr::rename("value" = item) %>%
  mutate(item = "ses_subjective")

ethnicity_self = demo_clean %>%
  filter(scale_name == "race_self" & !is.na(value)) %>%
  mutate(scale_name = "race_ethnicity_self",
         item = "race_ethnicity_self")

ethnicity = demo_clean %>%
  filter(scale_name == "race" & !is.na(value)) %>%
  mutate(item = dplyr::recode(item, "1" = "White",
                              "2" = "Black or African American",
                              "3" = "East Asian",
                              "4" = "American Indian or Alaskan Native",
                              "5" = "Native Hawaiian or Other Pacific Islander",
                              "6" = "South Asian",
                              "7" = "Middle Eastern or North African ",
                              "8" = "Prefer not to say",
                              "9" = "Racial/ethnic identity not listed"), item) %>%
  select(-value) %>%
  rename("item" = scale_name,
         "value" = item) %>%
  mutate(scale_name = "race_ethnicity",
         item = "race_ethnicity")

demo = demo_clean %>%
  filter(grepl("^age$|degree|household|hispanic_latinx|state|zip", item)) %>%
  mutate(value = ifelse(grepl("hispanic_latinx", item), dplyr::recode(value, "0" = "No", 
                                                                      "1" = "Yes",
                                                                      "2" = "Prefer not to say"), value),
         value = ifelse(grepl("degree", item), dplyr::recode(value, "1" = "No formal education",
                                                             "2" = "Did not complete high school",
                                                             "3" = "High school graduate (diploma)",
                                                             "4" = "High school graduate (GED)",
                                                             "5" = "Some college (1-4 years, no degree)",
                                                             "6" = "Associate's degree (including occupational or academic degrees)",
                                                             "7" = "Bachelor's degree (BA, BS, etc)",
                                                             "8" = "Master's degree (MA, MS, MENG, MSW, etc)",
                                                             "9" = "Professional school degree (MD, DDC, JD, etc)",
                                                             "10" = "Doctorate degree (PhD, EdD, etc)"), value),
         value = ifelse(grepl("household", item), dplyr::recode(value, "1" = "Less than $5,000",
                                                                "2" = "$5,000 through $11,999",
                                                                "3" = "$12,000 through $15,999",
                                                                "4" = "$16,000 through $24,999",
                                                                "5" = "$25,000 through $34,999",
                                                                "6" = "$35,000 through $49,999",
                                                                "7" = "$50,000 through $74,999",
                                                                "8" = "$75,000 through $99,999",
                                                                "9" = "$100,000 through $149,999",
                                                                "10" = "$150,000 and greater",
                                                                "11" = "Prefer not to say",
                                                                "12" = "NA"), value)) %>%
  bind_rows(., ses_subjective, ethnicity, ethnicity_self) %>%
  arrange(SID) %>%
  select(SID, condition, scale_name, item, value)
```

# write csvs
```{r}
write.csv(filter(daily, !SID %in% exclusions$SID), "../../Data/Prepped/daily.csv", row.names = FALSE)
write.csv(filter(dvs_scored, !SID %in% exclusions$SID), "../../Data/Prepped/dvs.csv", row.names = FALSE)
write.csv(filter(ind_diffs_merged, !SID %in% exclusions$SID), "../../Data/Prepped/individual_diffs.csv", row.names = FALSE)
write.csv(filter(intervention, !SID %in% exclusions$SID), "../../Data/Prepped/intervention.csv", row.names = FALSE)
write.csv(filter(feedback, !SID %in% exclusions$SID), "../../Data/Prepped/feedback.csv", row.names = FALSE)
write.csv(filter(demo, !SID %in% exclusions$SID), "../../Data/Prepped/demographics.csv", row.names = FALSE)
write.csv(exclusions, "../../Data/Prepped/exclusions.csv", row.names = FALSE)
```
