---
title: "WiN: data cleaning post"
author: "Dani Cosme & Taurean Butler"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE)
options(scipen=999)
```

# load packages
```{r}
if(!require('pacman')) {
	install.packages('pacman')
}

pacman::p_load(tidyverse, knitr, DT, devtools)

if (!require(scorequaltrics)) {
  devtools::install_github('dcosme/qualtrics', ref = "dev/enhance")
}
```

# define variables and paths

To pull data from Qualtrics, you need a credentials file with an API token associated with your account. To create the file, follow these steps.

1. Generate an API token for Qualtrics. Follow the steps outlined [here](https://www.qualtrics.com/support/integrations/api-integration/overview/).

2. Save a Qualtrics credentials text file with the following format. In this example, the file is being saved as `~/credentials.yaml.PENN`. The `baseurl` is the URL for your institution on Qualtrics. Use `upenn.co1.qualtrics.com` for Penn Qualtrics.

```
token: oILNW6...[your qualtrics API token]
baseurl: upenn.co1.qualtrics.com
```

`cred_file_location` = path to your Qualtrics credential file. 

`survey_name_filter` = regular expression to filter the available surveys

```{r}
survey_name_filter = "WiN End Survey - FINAL"
keep_columns = '(ResponseId|Finished|Progress|PROLIFIC_PID|ID|failed|condition|activeweek|StartDate)'
cred_file_location = "~/qualtrics_credentials.yaml"
#cred_file_location = "~/Documents/qualtrics_config.yaml"
ignore_items = "ResponseId|IPAddress|RecipientFirstName|RecipientLastName|RecipientEmail|ExternalReference|LocationLatitude|LocationLongitude"
```

# fetch qualtrics data
The Qualtrics API is pretty finicky. If you get the following error, just rerun the `get_survey_data` command until it works:

```
Error in qualtrics_response_codes(f, raw = TRUE) : 
  Qualtrics API complains that the requested resource cannot be found (404 error).
Please check if you are using the correct survey ID.
```
```{r}
# load data file if it exists, or pull fresh from qualtrics using the API
if (!file.exists("../../Data/Raw/post_raw.RDS")) {
  # load credential file
  credentials = scorequaltrics::creds_from_file(cred_file_location)
  
  # filter
  surveysAvail = scorequaltrics::get_surveys()
  surveysFiltered = filter(surveysAvail, grepl(survey_name_filter, name))
  knitr::kable(arrange(select(surveysFiltered, name), name))
  
  # fetch data
  surveys_long = scorequaltrics::get_survey_data(surveysFiltered,
                                                 credentials, 
                                                 pid_col = keep_columns) %>%
                 filter(!grepl(ignore_items, item)) %>% #filter out identifiable data
    select(survey_name, condition, activeweek, StartDate, ResponseId, PROLIFIC_PID, Finished, Progress, item, value)

  saveRDS(surveys_long, "../../Data/Raw/post_raw.RDS")
} else {
  surveys_long = readRDS("../../Data/Raw/post_raw.RDS")
}

```


# tidy

TO FIX

!! includes timers that include "_t_" rather than starting with "timing" --> rename in qualtrics after data is collected

```{r}
# load id key
id_key = read.csv("../../Data/Cleaned/pre_key.csv")

# filter out non-participant responses and recode IDs
surveys_included = surveys_long %>%
  filter(!is.na(PROLIFIC_PID)) %>% #remove test responses
  filter(!grepl("5efe5bdf6b4d873597a3c3e6", PROLIFIC_PID)) %>% #remove test responses
  filter(nchar(PROLIFIC_PID) == 24) %>% # remove invalid IDs from test responses
  filter(Progress > 90) %>% # remove incomplete responses 
  filter(!grepl("_DO_", item)) %>% # remove order items
  filter(!grepl("Q_Recaptcha|Q_Duplicate|^Status$|DistributionChannel|EndDate|RecordedDate|UserLanguage|Last Seen", item)) %>% #remove qualtrics variables we wont' use
  select(-Finished, -Progress) # update if we include people who completed less than 100%

# tidy scales with multiple items
surveys_subitems = surveys_included %>% 
  filter(grepl("_[0-9]+$", item)) %>%
  mutate(# slider numbers
         item = ifelse(grepl("alcohol_craving_1|alcohol_thoughts_1", item), gsub("_1", "", item), item)) %>% 
  extract(item, c("scale_name", "number"), "(.*)_([0-9]{1,2}$)", remove = FALSE) %>% # separate scale names and item numbers
  mutate(scale_name = ifelse(is.na(scale_name), item, scale_name),
         item = ifelse(!is.na(number), number, item)) %>%
  select(-number)

# timers
surveys_timer = surveys_included %>%
  filter(grepl("timing_|_t_", item)) %>% # !! includes timers that include "_t_" rather than starting with "timing" --> rename in qualtrics after data is collected
  filter(!grepl("_timing_", item)) %>%
  extract(item, c("item"), "timing_(.*)", remove = FALSE) %>%
  dplyr::mutate(scale_name = "timing")

# tidy remaining items and merge
surveys_tidy <- surveys_included %>% 
  dplyr::filter(!grepl("_[0-9]+$", item)) %>%
  dplyr::filter(!grepl("^timing_|_timing_|_t_", item)) %>% # !! includes timers that include "_t_" rather than starting with "timing" --> rename in qualtrics after data is collected
  tidyr::extract(
    item,
    c("scale_name", "sub_item"),
    "^(SNI_work|ux_feedback)_(.*)$",
    remove = FALSE
  ) %>%
  dplyr::mutate(
    scale_name = ifelse(is.na(scale_name), item, scale_name),
    item = ifelse(!is.na(sub_item), sub_item, item),
    item = gsub("8_", "", item), # remove extraneous number from open-ended item
    scale_name = gsub("8_", "", scale_name), # remove extraneous number from open-ended item
    item = gsub(".*_TEXT", "other", item), # recode open response items
    scale_name = gsub("_TEXT", "", scale_name), # recode open response items
    value = ifelse(value == "", NA, value),
  ) %>%
  dplyr::select(-sub_item) %>%
  dplyr::bind_rows(surveys_subitems) %>%
  dplyr::bind_rows(surveys_timer)

# check for duplicates and fix as needed
sid_duplicates = surveys_included %>%
  spread(item, value) %>%
  select(survey_name, ResponseId, StartDate, PROLIFIC_PID, condition) %>%
  unique() %>%
  group_by(PROLIFIC_PID) %>%
  dplyr::mutate(n = n()) %>%
  filter(n > 1) %>%
  arrange(PROLIFIC_PID, StartDate)

# create new SIDs
duplicate_ids = sid_duplicates %>%
  dplyr::mutate(second_survey = ifelse(!row_number() == 1, "yes", "no")) %>%
  filter(second_survey == "yes")

# merge
surveys_id = surveys_tidy %>%
  filter(!ResponseId %in% duplicate_ids$ResponseId) %>% # remove second of duplicate responses if applicable
  left_join(., id_key) %>%
  arrange(SID, scale_name, item) %>%
  select(survey_name, StartDate, SID, PROLIFIC_PID, scale_name, condition,  item, value) #removed "failed_attention" and  variables from original cleaning pre script because do not exist in surveys_included

# check id merging
# these two people are missing condition info and activeweek in qualtrics and therefore the data didn't merge correctly
missing_ids = surveys_id %>% filter(is.na(SID)) %>%
  left_join(., id_key, by = c("PROLIFIC_PID")) %>% 
  select(-contains(".x")) %>%
  rename("SID" = SID.y,
         "condition" = condition.y)

# fix missing condition info and ids
surveys  = surveys_id %>%
  filter(!PROLIFIC_PID %in% missing_ids$PROLIFIC_PID) %>%
  bind_rows(missing_ids) %>%
  arrange(SID, scale_name, item) %>%
  mutate(day = 13) %>% # add day variable
  select(survey_name, StartDate, condition, day, SID, scale_name, item, value)

```

# summarize survey ns {.tabset}
## number of scale response
```{r}
surveys %>%
  filter(!is.na(value)) %>%
  filter(!grepl("timing|Duration|User|Distribution|Date|Status|Task|perPage|time", scale_name)) %>%
  group_by(scale_name) %>%
  select(SID, scale_name) %>%
  unique() %>%
  dplyr::summarize(n = n()) %>%
  DT::datatable(filter = "top", rownames = FALSE)
```

## number of scale responses per person
```{r}
surveys %>%
  filter(!is.na(value)) %>%
  filter(!grepl("timing|Duration|User|Distribution|Date|Status|Task|perPage|time", scale_name)) %>%
  group_by(SID, scale_name) %>%
  summarize(n_responses = n()) %>%
  group_by(scale_name, n_responses) %>%
  summarize(n_people = n()) %>%
  DT::datatable(filter = "top", rownames = FALSE)
```

## number of missing responses
```{r}
surveys %>%
  filter(is.na(value)) %>%
  filter(!grepl("timing|Duration|User|Distribution|Date|Status|Task|perPage|time|pageNo", scale_name)) %>%
  group_by(SID, scale_name, item) %>%
  group_by(scale_name, item) %>%
  summarize(n_people_missing = n()) %>%
  DT::datatable(filter = "top", rownames = FALSE)
```

## number of participants {.tabset}
### all 
```{r}
surveys %>%
  select(SID) %>%
  unique() %>%
  dplyr::summarize(n = n())
```

### by condition 
```{r}
surveys %>%
  select(SID, condition) %>%
  unique() %>%
  group_by(condition) %>%
  dplyr::summarize(n = n())
```

# check scale ranges
```{r}
surveys %>%
  dplyr::filter(
    !grepl(
      "timing|Duration|User|Distribution|Date|Status|Task|perPage|time|zipcode|military_date|activeweek|Recorded|EndDate|StartDate|consent|worktimeArray",
      scale_name
    )
  ) %>%
  dplyr::mutate(
    value_num = suppressWarnings(as.numeric(value))
  ) %>%
  dplyr::filter(!is.na(value_num)) %>%
  dplyr::group_by(scale_name) %>%
  dplyr::summarise(
    min = round(min(value_num), 2),
    max = round(max(value_num), 2),
    mean = round(mean(value_num), 2),
    var = round(var(value_num), 2),
    .groups = "drop"
  ) %>%
  dplyr::arrange(scale_name) %>%
  DT::datatable(filter = "top", rownames = FALSE)
```

# write csvs
```{r}
write.csv(surveys, "../../data/cleaned/post_clean_long.csv", row.names = FALSE)
```

